---
title: "EDA_Reworked"
output: html_document
date: "2023-04-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

# Import Data Below
steam_games <- read.csv("steam.csv")

#function for finding midpoint of owners and converting to numeric
find_mids = function(owners) {
  rowMid <- rowMeans(matrix(as.numeric(unlist(str_split(owners, pattern = "-"))),ncol = 2, byrow = T))
  return(rowMid)
}

steam_games_final = steam_games %>%
  mutate(mid_owner=find_mids(owners)) %>%
  mutate(revenue = mid_owner*price) %>%
  rename(age=required_age) %>%
  mutate(num_platforms = str_count(platforms,";")+1, .after=platforms) %>%
  filter(english==1) %>%
  mutate(release_year=str_replace_all(release_date,"-", "")) %>%
  mutate(release_year=substr(release_year,1,4)) %>%
  mutate(pos_rate_prop=positive_ratings/(positive_ratings+negative_ratings))%>%
  mutate(neg_rate_prop=negative_ratings/(positive_ratings+negative_ratings)) %>%
  filter()

steam_games_final$release_year <- as.numeric(as.character(steam_games_final$release_year))
#steam_games_final

desired_genres <- c("Indie","Action","Casual","Adventure","Strategy","Simulation","RPG","Early Access","Free to Play","Sports")

steam_games_genre <- subset(steam_games_final, genres %in% desired_genres) #we can use this dataset for modeling with genre because it gets rid of multiple genre categories?

#steam_games_genre_fill <- separate_rows(steam_games_final, genres, sep = ";")
```

# Creator: Sophia Palmieri

### Q1: Are positive ratings higher for games that have longer average playtimes?

```{r}
steam_games_final=steam_games_final%>%
  mutate(pos_rate_prop=positive_ratings/(positive_ratings+negative_ratings))%>%
  mutate(neg_rate_prop=negative_ratings/(positive_ratings+negative_ratings))

playtimev.ratings= steam_games_final %>% 
  select(pos_rate_prop, average_playtime,developer)%>%
  filter(average_playtime<75000)%>%
  filter(average_playtime>0)
  
ggplot(playtimev.ratings,aes(x=average_playtime,y=pos_rate_prop))+geom_point()+labs(x="Average Playtime", y="Proportion of Potitive Ratings", title="Positive Ratings vs. Average Playtime")

##I filtered out all average playtimes that were 0 because games have to be played to be rated and greater than 75000 to reduce outliers. 
```

### Q2: Is there a correlation between the number of owners and the price?

```{r}
find_mids = function(owners) {
  rowMid <- rowMeans(matrix(as.numeric(unlist(str_split(owners, pattern = "-"))),ncol = 2, byrow = T))
  return(rowMid / 1000)
}

ownersv.price= steam_games_final %>% 
  select(price,owners) %>%
  filter(owners!=0) %>%
  group_by(price) %>%
  summarise(mean_owners = mean(find_mids(owners)))


ggplot(ownersv.price,aes(x=price, y=mean_owners)) + 
  geom_point(color="red") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + labs(x="Price ($)", y="Number of Owners (Thousands)",title="Price vs. Number of Owners (In Thousands)") +
  scale_y_log10() +
  scale_x_log10()

```

# Interpreter: Alison Polito

### Q1: Is there a relationship between the release date of a game and its age requirements? 

```{r}
steam_games_1 = steam_games_final %>%
  select(release_date,age) %>%
  mutate(release_year=str_replace_all(release_date,"-", "")) %>%
  mutate(release_year=substr(release_year,1,4)) %>%
  group_by(release_year) %>%
  summarize(mean_age=mean(age)) 

ggplot(steam_games_1, aes(x=as.numeric(release_year),y=mean_age)) + 
  geom_point() + geom_smooth(method=lm,se=FALSE) + xlab("Release Year") + ylab("Average Required Age") + ggtitle("Release Year vs. Average Required Age to Play") 


# Note: many with 0 are unrated or unsupplied.
# group_by month and year
```

### Q2: Is there a relationship between the number of platforms that the game can be played on and number of owners?

```{r}
find_mids = function(owners) {
  rowMid <- rowMeans(matrix(as.numeric(unlist(str_split(owners, pattern = "-"))),ncol = 2, byrow = T))
  return(rowMid / 1000)
}

#owner_test = c("100-200", "200-400")
#test1 <- find_mids(owner_test)
#head(test1)

#test <- sapply(steam_games_final$owners, find_mids)
#head(test)

steam_games_4 = steam_games_final %>%
  mutate(mid_owner=find_mids(owners)) %>%
  group_by(num_platforms) %>%
  summarize(mean_mid_owner = mean(mid_owner))
steam_games_4

```

# Interpreter: Srinidhi Boyanapalli 

### Q1: Are some genres more expensive than others?

```{r}
team_games_genre <- steam_games_final %>%
  select(genres, price) %>%
    group_by(genres) %>%
  summarise(avg_price=mean(price)) %>%
    arrange(desc(avg_price)) 

head(team_games_genre)

#It looks like with Animation and Modeling and Design having the highest average price, some genres are indeed more expensive than the others presumably because of the demand of these genres.
```

### Q2: What is the relationship between the price of the product and the median playtime?

```{r}
steam_games_price_median_playtime = steam_games_final %>% 
  select(price, median_playtime)
ggplot(steam_games_final,aes(x= price, y= median_playtime))+geom_point()

#It looks like the median playtime mostly exists under 50000 and is high for prices that are lower than 100. So, cheaper games most likely have a greater median playtime presumably because of the affordability and accessibility.
```

# Deliverer: Neha Palle

### Q1: Is there a correlation between the latest update and the number of negative ratings for a game?

```{r}
co5 = steam_games_final %>%
  select(release_date,negative_ratings) %>%
  mutate(release_date=str_replace_all(release_date,"-", "")) %>%
  mutate(release_year=substr(release_date,1,4),release_month=substr(release_date,5,6))


co5_graph = co5 %>%
  ggplot(aes(x=release_year,y=negative_ratings)) +
  geom_bar(stat="identity",fill="pink") +
  xlab("Release Year") + ylab("Number of Negative Ratings") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + labs(title = "Release Year vs. Number of Negative Ratings")

co5_graph
```

### Q2: Is there a correlation between required age to play the game and the number of achievements? 

```{r}
co6 = steam_games_final %>%
  group_by(age) %>%
  summarise(average_achievements = mean(achievements))

ggplot(co6,aes(x=as.character(age),y=average_achievements)) +
  geom_bar(stat='identity',fill="pink") +
  xlab("Age Required to Play") + ylab("Average Achievements") +
  theme_minimal() + 
  labs(title="Age Required vs. Average Achievements")
```

# Orator: Truc Pham

### Q1: Which developer has the most negative ratings?

```{r}
steam_games_negative <- steam_games_final %>% 
  select("developer", "negative_ratings") %>%
  group_by(developer) %>% 
  summarise_each(funs(sum)) %>%
  arrange(desc(negative_ratings))

head(steam_games_negative, 5)
```

### Q2: How does the number of in-game achievements correlate with the median playtime? 

```{r}
achievements <- steam_games_final %>%
  select(categories, achievements, median_playtime) %>%
  group_by(categories) %>% 
  summarise_each(funs(sum)) %>%
  arrange(desc(achievements))

achievements 

ggplot(steam_games_final,aes(x=median_playtime,y=achievements))+geom_point() + geom_smooth() +
  ggtitle("In-Game Achievements vs. median Playtime")
```

# Orator: Girija Joshi

### Q1: Is there a relationship between the amount of in game achievements and positive ratings? 

```{r}
q11 <- steam_games %>% 
  select(achievements, positive_ratings, negative_ratings) %>%
  mutate(prop_pos = positive_ratings/(positive_ratings+negative_ratings)) %>%
  select(achievements, prop_pos) %>% 
  filter(achievements<8000)

head(q11)

q11_graph <- q11 %>%
  ggplot(aes(x=achievements, y=prop_pos)) + 
  geom_point() + 
  geom_smooth()+
  xlab("#of In-Game Achievements") + ylab("Proportion of Positive Ratings") +
  ggtitle("How The Proportion of Positive Ratings Changes According To \nthe # of In Game Achievements " )

q11_graph
```

### Q2: Are games with more positive ratings priced higher than those with negative ratings?

```{r}
q12 <- steam_games %>% 
  select(price, positive_ratings, negative_ratings) %>%
  mutate(prop_pos = positive_ratings/(positive_ratings+negative_ratings)) %>%
  select(price, prop_pos)

head(arrange(q12, desc(prop_pos)))


q12_graph <- q12 %>%
  ggplot(aes(x=prop_pos, y=price)) + 
  geom_point(aes(color = ifelse(prop_pos>0.5, 'salmon', 'turquoise'))) +
  scale_colour_manual(labels = c("More (-) Ratings", "More (+) Ratings"), values=c('turquoise', 'salmon')) + 
  labs(color = "Ratings")+
  xlab("Proportion of Positive Ratings") + ylab("Price($)") + 
  ggtitle("How Price Changes According to Rating")

q12_graph
```

# Follow-up Questions

### New Questions Based Off Initial Investigation

- Q1: How do the top ten developer games change in average playtime as their proportion of positive ratings increase?
- Q2: How does the price/proportion of negative ratings change across month and year?
- Q3: How do the number of owners for the most popular genres (top ten) change across time (years)? 
- Q4: Does the relationship between number of in-game achievements and positive ratings change over time? If so, how? 

### Investigation of Follow-up Questions

GIVE WHAT 2 QUESTIONS YOU ATTEMPTED TO INVESTIGATE FURTHER IN COMPLETE SENTENCES 

We decided to investigate Q1 and Q2 in further detail.

```{r}

DATA2 = na.omit(steam_games_final) %>%
  crossv_kfold(20)

```


SHOW AT LEAST 2 TABLES OR FIGURES BELOW THAT EXPLORE ANSWERS FOR THE QUESTIONS YOU ARE INVESTIGATING FURTHER.

```{r}

popular <- steam_games_final %>% 
   separate_rows(genres,sep=";") %>% 
   count(genres) %>%
   arrange(desc(n))

#Top ten genres = Indie, Action, Casual, Adventure, Strategy, Simulation, RPG, Early Access, Free to Play, Sports

```



```{r}
#How do the top ten developer games change in average playtime as their proportion of positive ratings increase?

develop= steam_games_final%>%
  select(positive_ratings, average_playtime, developer)%>%
  group_by(developer)%>%
  summarize(n=n())%>%
  arrange(desc(n))%>%
  head(10)
devplay=playtimev.ratings%>%
  semi_join(develop,by="developer")%>%
  filter(average_playtime<25000)

ggplot(devplay,aes(x=average_playtime,y=pos_rate_prop,color=developer)) +
  geom_point() +
  labs(x="Average Playtime", y="Proportion of Potitive Ratings", title="Positive Ratings vs. Average Playtime for Developers", color="Developers")

##Top 10 developers are shown here, but only 7 appear because I filtered out any average playtimes that were 0. I also filtered out any average playtimes over 25000 hours to remove a singular outlier.
```

```{r}
#How does the price/proportion of negative ratings change across month and year?


price_map = steam_games_final %>%
  select(release_date,negative_ratings,price) %>%
  mutate(release_date=str_replace_all(release_date,"-", "")) %>%
  mutate(release_year=substr(release_date,1,4),release_month=substr(release_date,5,6)) %>%
  group_by(release_year,release_month) %>%
  summarise(mean_price=mean(price))
  

ggplot(price_map,mapping = aes(x=release_month,y=release_year)) + geom_tile(mapping=aes(fill=mean_price)) + scale_fill_distiller(palette = "YlGn", direction=2) +
  labs(x="Release Month",y="Release Year",title = "Proportion of Price Across Month and Year")

nrating_map = steam_games_final %>%
  select(release_date,negative_ratings,positive_ratings)%>%
  mutate(release_date=str_replace_all(release_date,"-", "")) %>%
  mutate(release_year=substr(release_date,1,4),release_month=substr(release_date,5,6)) %>%
  group_by(release_year,release_month) %>%
  summarise(prop_neg_ratings = mean(negative_ratings/(negative_ratings + positive_ratings)))

ggplot(nrating_map,mapping = aes(x=release_month,y=release_year)) + geom_tile(mapping=aes(fill=prop_neg_ratings)) + scale_fill_distiller(palette = "YlGn", direction=2) + 
  labs(x="Release Month",y="Release Year",title = "Proportion of Negative Ratings Across Month and Year")
```
